---
title: "HR Attrition Exploratory Data Analyiss"
output: html_notebook
---

Loading the dataset

```{r}
library(ggplot2)
library(gridExtra)
library(grid)

hr = read.csv('HR_Employee_Attrition.csv')
dim(hr)
names(hr)
```
So about 35 columns, 1470 rows. The dependent variable will likely be "attrition" with everything else being the x - predictors. Let's explore the data a bit at first with a few questions:

- What are the summary statistics for attrition? On average, how long does it take for someone to leave the company? What's the distribution of this dependent variable look like?
- How do we define attrition? What metric are we using to measure this? 
- From my own intuition, which predictor variables do I think will have the highest prediction power, so I can test?
- Which variables are correlated with each other? 
- What are some qualitative predictor variables? What's does the distribution for these qualitative look like? Could there be subjective error there?

```{r}
summary(hr$Attrition)
```
```{r}
head(hr, 10)
```
Note: 
- To build the model, I'm gonna need to replace these categorical variables with dummy variables. 
- The class is fairly imbalanced so I'll need to keep this in mind when I go into the modeling stage. 

Let's look at multi-collinearity to see how related these quantitative variables are, so we have a solid first guess on which variables to test: 
```{r}
install.packages("pheatmap")
install.packages("RColorBrewer")
install.packages("viridis")
```
```{r}
library(pheatmap)
library(RColorBrewer)
library(viridis)
names(hr)
str(hr)

group_col = c('Age','DailyRate','DistanceFromHome','Education','EmployeeCount','EmployeeNumber','EnvironmentSatisfaction','JobInvolvement','JobLevel','JobSatisfaction','MonthlyIncome','MonthlyRate','NumCompaniesWorked','PercentSalaryHike','PerformanceRating','RelationshipSatisfaction','StandardHours','StockOptionLevel','TotalWorkingYears','TrainingTimesLastYear','WorkLifeBalance','YearsAtCompany','YearsInCurrentRole','YearsSinceLastPromotion','YearsWithCurrManager')

quant_var = hr[group_col]



pheatmap(
  mat               = quant_var,
  color             = inferno(10),
  border_color      = NA,
  show_colnames     = FALSE,
  show_rownames     = FALSE,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 14,
  main              = "Default Heatmap"
)

install.packages('ggplot2')

library('ggplot2')
library('reshape2')

cor_matrix = cor(quant_var)

melted_cormat = melt(cor_matrix)

ggplot(data = melted_cormat,aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

So it looks like the only four with a good amount of multi-collinearity is the "YearsWith" 4 categories. 

```{r}
head(hr)
summary(hr$Attrition)

names(hr)
```

```{r}
head(hr)

# Split dataset between employees turned over vs. employees still there
turned_over = subset(hr,Attrition == 'Yes')
incumbent = subset(hr, Attrition == 'No' )

p1 = ggplot(data = turned_over, aes(x = Age)) + 
  geom_histogram(binwidth = 1)

p2 = ggplot(data = incumbent, aes(x = Age)) + 
  geom_histogram(binwidth = 1)

grid.arrange(p1, p2)



```
```{r}
summary(turned_over$Age)

summary(incumbent$Age)
```
So it looks like on average, employees who stay tend to be older while employees who are younger tend to leave. 

```{r}
names(hr)

by(hr$JobLevel, hr$Attrition, summary)

ggplot(aes(x = hr$JobLevel), data = hr) + 
  geom_histogram(bins = 5)
```
Notes: It looks like most employees are at a lower joblevel, which makes sense, however, it looks like those who are at a lower job level are more likely to leave, my guess, is that it's because they're not getting promoted. 

```{r}
names(hr)
by(hr$PercentSalaryHike, hr$Attrition, summary)

p1 = ggplot(data = hr, aes(y = PercentSalaryHike, x = Attrition)) + 
  geom_boxplot() + 
  ggtitle("Percent Salary vs. Attrition")

p1

```
```{r}
names(hr)

hr$PerformanceRating

ggplot(data = hr, aes(x = PerformanceRating)) +
  geom_histogram(binwidth=0.5)
```
We can see that a majority of people get a rating of 3, while only a select few get a rating of 4. Let's look at the Performance Rating distribution from turned over employees versus non-turnover.

```{r}
ggplot(data = turned_over, aes(x = PerformanceRating)) +
  geom_histogram(binwidth=0.5)
```

```{r}
ggplot(data = incumbent, aes(x = PerformanceRating)) +
  geom_histogram(binwidth=0.5)

hr$Gender
```
You can see that there doesn't seem to be a big difference. 

```{r}
by(hr$Attrition,hr$Gender, summary)

```
Interesting. Looks men have about 3% higher turnover than women. Could there be a gender behavioral difference? 
```{r}
87/501
150/732
```
Let's look at business areas/departments now.
```{r}

hr$Attrition = as.numeric(factor(hr$Attrition, levels = c('Yes','No'))) 

aggregate(hr, by = list(hr$Department), FUN = mean)

hr$Attrition

```

